Estimating the dynamics of\\SPARQL query results over Wikidata\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}

### 2019-01-27 kgswc19.tex
Finished first pass over the entire paper

\section{Datasets}
\label{sec:data}

@@

\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured weekly in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property. The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%). Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the weeks progress, while on the right, shows the numbers of triples added and removed week-to-week. We see that although many more triples are added than removed, there are some triples removed each week. We further note some peaks in triples added in some weeks, which may be due to bulk imports of data. The gap in week 12 is due to the fact that we were not able to obtain the data for that week (we thus exclude the triples added/removed from weeks 11 to 12, and 12 to 13).

@@

\begin{figure}[t]
	%\begin{minipage}[b]{0.47\linewidth}
		\centering
		\begin{tikzpicture}[thick,font=\scriptsize]
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Version},
		width=5cm, 
		height=5cm,
		ymin=0,
		xtick={1,2,...,23},
		xticklabels={},
		extra x ticks={1,23},
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(13,1348622441)(14,1418460320)(15,1432143355)(16,1470298276)(17,1506140622)(18,1542278452)(19,1593334993)(20,1613832004)(21,1684774955)(22,1771601730)(23,1843086909)(24,1924967162)
		};

@@

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,115)(2,112)(3,118)(4,116)(5,121)(6,111)(7,127)(8,109)(9,118)(10,115)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(12,105)(13,113)(14,118)(15,117)(16,111)(17,118)(18,112)(19,116)(20,106)(21,106)(22,129)
		};

@@

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,21)(2,21)(3,21)(4,21)(5,21)(6,21)(7,21)(8,21)(9,21)(10,21)(11,21)(12,21)(13,21)(14,21)(15,21)(16,21)(17,21)(18,21)(19,21)(20,21)(21,21)(22,21)(23,21)(24,21)(25,21)(26,21)(27,21)(28,21)(29,21)(30,21)(31,21)(32,21)(33,21)(34,21)(35,21)(36,21)(37,21)(38,21)(39,21)(40,21)(41,21)(42,21)(43,21)(44,21)(45,20)(46,20)(47,20)(48,20)(49,20)(50,20)(51,20)(52,19)(53,19)(54,19)(55,19)(56,19)(57,19)(58,19)(59,19)(60,18)(61,18)(62,18)(63,18)(64,18)(65,18)(66,18)(67,17)(68,17)(69,17)(70,17)(71,17)(72,17)(73,17)(74,16)(75,16)(76,16)(77,16)(78,16)(79,16)(80,15)(81,15)(82,15)(83,15)(84,14)(85,14)(86,14)(87,14)(88,14)(89,14)(90,14)(91,14)(92,14)(93,13)(94,13)(95,13)(96,13)(97,13)(98,13)(99,12)(100,12)(101,12)(102,12)(103,12)(104,12)(105,12)(106,12)(107,12)(108,12)(109,11)(110,11)(111,11)(112,11)(113,10)(114,10)(115,10)(116,10)(117,10)(118,9)(119,9)(120,9)(121,9)(122,9)(123,9)(124,9)(125,9)(126,8)(127,8)(128,8)(129,8)(130,8)(131,8)(132,8)(133,8)(134,8)(135,8)(136,7)(137,7)(138,7)(139,7)(140,7)(141,7)(142,7)(143,6)(144,6)(145,6)(146,6)(147,6)(148,6)(149,6)(150,6)(151,5)(152,5)(153,5)(154,4)(155,4)(156,4)(157,4)(158,4)(159,4)(160,4)(161,4)(162,4)(163,4)(164,4)(165,4)(166,3)(167,3)(168,3)(169,3)(170,3)(171,3)(172,2)(173,2)(174,2)(175,2)(176,2)(177,2)(178,2)(179,2)(180,2)(181,2)(182,2)(183,2)(184,2)(185,2)(186,2)(187,2)(188,1)(189,1)(190,1)(191,1)(192,1)(193,1)(194,1)(195,1)(196,1)(197,1)(198,1)(199,1)(200,1)(201,1)(202,1)(203,1)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
		};

@@

@@

We evaluate the performance of our proposal to predict changes in the results of our queries. For this, we use the Wikidata dataset and our query dataset explained previously.
To build the Difference Sets ($\Delta$) for our model we used an approach based on a merge-sort to scan over the snapshots as follows:

\begin{enumerate}
	\item Sort all statements by their syntactic order (subject-predicate-object).
	\item Perform a pairwise comparison of the statements by scanning two snapshots in linear time.
	\item Trigger a detection of the change as soon as the order of the statements differs between two snapshots.
\end{enumerate}

Then we configure the parameters of our model. For the time interval, we consider three versions of the dataset equivalent to the last three weeks. As the Transformation Function, we evaluate three basic functions: Maximum, Minimum, and Average. For example, Maximum will take the dynamics of the most dynamic property (used as a predicate in the query) as a feature.
In addition, given that existing approaches are based on the historical results of the queries, we evaluate the substitution of our model for a historical change indicator that counts how many times the results have changed in the last three weeks.

In addition, we train a Machine Learning model to determine the life spam of a query results. To train the model we consider the changes observed in the monitored interval. For each update of the results of a query, we have created an example of training with the previously presented features. The target TTL value that we learn is the time interval during which the results of the query remain fresh from the current update time.

\subsection{Results}

First, we consider the Pearson correlation between the characteristics. We found only a couple of highly correlated variables (\# triples and \# vars: 0.943). The influence of the use of various Transformation Functions in our model was also evaluated. In this case, Average was the Transformation Function that obtained the best result followed by Maximum.

We tested a list of classifiers, including Decision Tree, Nearest Neighbors, and Naive Bayes among others listed in Table~\ref{tab:classifier}. Our pilot experiment shows that Nearest Neighbors and Decision Tree work better than the others. We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed features. We split the data 80\% for training and 20\% for tests. To avoid overfitting, we use cross-validation in the training data.

Table~\ref{tab:classifier} shows the quality values of the classifiers evaluated in terms of $F_1$. To have a measure of the positioning of our proposal, we added a new feature that counts how many times the results of the query change on average in the last three versions. Then, we did two executions with our syntactic features: one using this new indicator, and the second with our dynamic model. Table~\ref{tab:classifier} shows that our algorithm achieved less quality. Despite obtaining lower results, the quality of our proposal is not far from that achieved by observing the results of queries in the past, but certainly, our approach is less expensive since it does not require knowledge of past query results.

Finally, to predict the time that the results of a query will remain invariant (TTL), a continuous value (number of weeks) will be determined. For that we employed techniques of multiple linear regression using the sklearn and tensorflow library. The latter allows us to see how the periods of estimation and the convergence analysis will behave.

We were able to predict the time of life (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques.

\begin{table*}[h]
	\centering
	\caption{Tested classifiers.}
	\label{tab:classifier}
	\begin{tabular}{|l|r|r|}    \hline		
		Classifier                   & Past Results & Dynamics Model \\    \hline
		Base Dummy 					 & 0.497869 & 0.486923 \\
		Nearest Neighbors            & 0.701587 & 0.635923 \\
		Decision Tree                & 0.695591 & 0.651462 \\
		Linear SVM                   & 0.705755 & 0.593846 \\
		Naive Bayes                  & 0.661239 & 0.587692 \\    \hline
	\end{tabular}
\end{table*}
\begin{comment}

@@

In this paper we propose classifiers that predict whether or not the results of a query will change in the short time interval; for this, we propose a model based on the dynamics of the predicates in the datasets and a set of features extracted from the syntax of the queries. The proposed framework is flexible for incorporating new features extracted from the queries, from the current results, or from the model based on the multiplicity of predicates.

We select the datasets to evaluate our proposal considering desirable characteristics, such as availability, size, dynamics, and integrity. Choosing Wikidata, we present statistics to understand such characteristics. In the case of the queries dataset, we did not find any existing dataset of queries that complied with the requirements, so we curated a query dataset for Wikidata based on query examples proposed by users.

Finally, we evaluate our proposal, making a weekly prediction of change in the results of the queries and evaluating the success rates of the different classifiers each week based on F-measure. We also apply linear regression techniques to predict, specifically, when the results of a query will change.
We measure the effectiveness of the characteristic based on our model of predicate dynamics versus a characterization of dynamics based on the historical results. The latter is a common factor in the previous proposals, but it requires a high cost. Our results show that our approach using predicate dynamics is competitive for prediction while being less expensive to support.

Although the results obtained are encouraging, they are still preliminary. We consider for future work evaluating the quality of our proposal over a longer period. In addition, we believe that more complex query features can improve the results, such as considering operator semantics; for example, to see if negation is affecting predicates or if the predicates are affected by the existence of filters. On the other hand, it would be worth expanding the set of queries in the test dataset, in order to increase the size, complexity, and coverage. Recently a query dataset for Wikidata was published\footnote{https://lists.w3.org/Archives/Public/semantic-web/2018Aug/0023.html}. It would be good to evaluate this query dataset, as well as consider a probabilistic approach based on query algebra (rather than syntax).

>>>>

\section{Gold Standard Dataset}
\label{sec:data}

@@

\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured weekly in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property. The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%). Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the weeks progress, while on the right, shows the numbers of triples added and removed week-to-week. We see that although many more triples are added than removed, there are some triples removed each week. We further note some peaks in triples added in some weeks, which may be due to bulk imports of data. The gap in week 12 is due to the fact that we were not able to obtain the data for that week (we thus exclude the triples added/removed from weeks 11 to 12, and 12 to 13). The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.

@@

\begin{figure}[t]
	%\begin{minipage}[b]{0.47\linewidth}
		\centering
		\begin{tikzpicture}[thick,font=\scriptsize]
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Version},
		width=5cm, 
		height=5cm,
		ymin=0,
		xtick={1,2,...,24},
		xticklabels={},
		extra x ticks={1,24},
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=blue,
		mark=diamond,
		]
		coordinates {
			(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)
		};

		\addplot[
		color=blue,
		mark=diamond,
		]
		coordinates {
			(13,1348622441)(14,1418460320)(15,1432143355)(16,1470298276)(17,1506140622)(18,1542278452)(19,1593334993)(20,1613832004)(21,1684774955)(22,1771601730)(23,1843086909)(24,1924967162)
		};

@@

		\addplot[
		color=blue,
		mark=diamond,
		]
		coordinates {
			(1,115)(2,112)(3,118)(4,116)(5,121)(6,111)(7,127)(8,109)(9,118)(10,115)
		};

		\addplot[
		color=blue,
		mark=diamond,
		]
		coordinates {
			(12,105)(13,113)(14,118)(15,117)(16,111)(17,118)(18,112)(19,116)(20,106)(21,106)(22,129)
		};

@@

		\addplot[
		color=blue
		]
		coordinates {
			(1,21)(2,21)(3,21)(4,21)(5,21)(6,21)(7,21)(8,21)(9,21)(10,21)(11,21)(12,21)(13,21)(14,21)(15,21)(16,21)(17,21)(18,21)(19,21)(20,21)(21,21)(22,21)(23,21)(24,21)(25,21)(26,21)(27,21)(28,21)(29,21)(30,21)(31,21)(32,21)(33,21)(34,21)(35,21)(36,21)(37,21)(38,21)(39,21)(40,21)(41,21)(42,21)(43,21)(44,21)(45,20)(46,20)(47,20)(48,20)(49,20)(50,20)(51,20)(52,19)(53,19)(54,19)(55,19)(56,19)(57,19)(58,19)(59,19)(60,18)(61,18)(62,18)(63,18)(64,18)(65,18)(66,18)(67,17)(68,17)(69,17)(70,17)(71,17)(72,17)(73,17)(74,16)(75,16)(76,16)(77,16)(78,16)(79,16)(80,15)(81,15)(82,15)(83,15)(84,14)(85,14)(86,14)(87,14)(88,14)(89,14)(90,14)(91,14)(92,14)(93,13)(94,13)(95,13)(96,13)(97,13)(98,13)(99,12)(100,12)(101,12)(102,12)(103,12)(104,12)(105,12)(106,12)(107,12)(108,12)(109,11)(110,11)(111,11)(112,11)(113,10)(114,10)(115,10)(116,10)(117,10)(118,9)(119,9)(120,9)(121,9)(122,9)(123,9)(124,9)(125,9)(126,8)(127,8)(128,8)(129,8)(130,8)(131,8)(132,8)(133,8)(134,8)(135,8)(136,7)(137,7)(138,7)(139,7)(140,7)(141,7)(142,7)(143,6)(144,6)(145,6)(146,6)(147,6)(148,6)(149,6)(150,6)(151,5)(152,5)(153,5)(154,4)(155,4)(156,4)(157,4)(158,4)(159,4)(160,4)(161,4)(162,4)(163,4)(164,4)(165,4)(166,3)(167,3)(168,3)(169,3)(170,3)(171,3)(172,2)(173,2)(174,2)(175,2)(176,2)(177,2)(178,2)(179,2)(180,2)(181,2)(182,2)(183,2)(184,2)(185,2)(186,2)(187,2)(188,1)(189,1)(190,1)(191,1)(192,1)(193,1)(194,1)(195,1)(196,1)(197,1)(198,1)(199,1)(200,1)(201,1)(202,1)(203,1)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
		};

@@

\paragraph{Feature extraction:} To compute the query features, we use Apache Jena to parse the query and extract the necessary statistics and determine the presence of the relevant query operators. In order to extract statistics on the dynamics of individual predicates in the dynamic graph, a challenge here is scalability, since we work with a total of 32.3 billion triples; hence we (1) sort each version of Wikidata, (2) apply a merge-sort iterator over each pair of consecutive versions ($G_i$ and $G_{i+1}$) to detect triples that changed ($G_i \oplus G_j$), writing a separate file for triples that were deleted ($G_i - G_j$) and added ($G_j - G_i$), (3) from these files, we can then compute and sum the number of triples changed for each predicate between each pair of versions. Finally, in order to create the ground truth in terms of which results change between which versions, we index each version in Virtuoso and compute the results for each query against each version and write them to disk; we then compare consecutive pairs of results with a merge-sort.

@@

%We evaluate the performance of our proposal to predict changes in the results of our queries. For this, we use the Wikidata dataset and our query dataset explained previously. To build the Difference Sets ($\Delta$) for our model we used an approach based on a merge-sort to scan over the snapshots as follows:
%
%\begin{enumerate}
%	\item Sort all statements by their syntactic order (subject-predicate-object).
%	\item Perform a pairwise comparison of the statements by scanning two snapshots in linear time.
%	\item Trigger a detection of the change as soon as the order of the statements differs between two snapshots.
%\end{enumerate}

Our experiments aim to ascertain the (relative) quality of predictions that can be made in the context of One Shot Change (OSC) and Time To Live (TTL) based on query features, statistics of predicate dynamics, as well as information about the historical evolution of queries.  \ah{Update with research questions later}

\paragraph{Setting:} Given that our tasks are prediction based, we must thus build a training set and a test set. With 23 versions, we must hold out at least one version for testing, leaving a maximum interval of 22 past versions to use for predictions. However, the more versions we add to interval of past versions considered, the fewer examples we can generate; for example, if we consider 22 versions for training, we will only have one ground truth label for each query. In the end, we opted to consider the 3 previous versions ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in total. We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted. (\ah{Discuss the gap: \textit{We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed features.}}) \ah{Could we do experiments with different intervals?}

In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates. In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.

We tested the Pearson correlation ($\rho$) between the features available in our data. We found only one pair of highly correlated variables: \textbf{\textnumero\ of triple patterns} and \textbf{\textnumero\ of variables} ($\rho = 0.943$). \ah{Not sure about this part: what does it add? What does ``highly correlated mean?'' Also it's surprising that, e.g., \textbf{\textnumero of predicates} is not also highly correlated here.}

For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes and Nearest Neighbors. We split the data to use 80\% for training and 20\% for tests. To avoid overfitting, we use cross-validation in the training data. We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) and information on historical results (\texttt{r}). \ah{We should be more explicit on precisely what \textsc{r} is?} We may also consider combinations of these features: \textsc{qp}, \textsc{qs}, \textsc{pr} and \textsc{qpr}. In our experiments, we compare \textsc{qp} with \textsc{r}. \ah{Is this correct? Or would it be \textsc{qr}?} \ah{Why only 2? Could we not compare all seven?}

%In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates. In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.

%Then we configure the parameters of our model. For the time interval, we consider three versions of the dataset equivalent to the last three weeks. As the Transformation Function, we evaluate three basic functions: Maximum, Minimum, and Average. For example, Maximum will take the dynamics of the most dynamic property (used as a predicate in the query) as a feature.
%In addition, given that existing approaches are based on the historical results of the queries, we evaluate the substitution of our model for a historical change indicator that counts how many times the results have changed in the last three weeks.

In addition, we train a Machine Learning model to determine the TTL of a query's results. To train the model we consider the changes observed in the monitored interval. For each update of the results of a query, we have created an example of training with the previously presented features. The target TTL value that we learn is the time interval during which the results of the query remain fresh from the current update time. \ah{This paragraph is a little vague/unclear to me.}

\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-measure for the predictions made; for reference, we include a baseline that randomly guesses yes/no. The results show that all configurations clearly outperform the random baseline; however, with the highest result being an $F_1$-measure of 0.706, we can also see that predicting changes in query results is indeed a challenging problem. Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available. However, the \textsc{qp} features -- which do not assume the availability of such information -- are often competitive: more specifically, the best result for \textsc{r} is 0.706 with Linear SVM, while the best results for \textsc{qp} is 0.651 with Decision Trees.

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} and \textsc{tensorflow} libraries. The latter allows us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques. \ah{Could do with a comparison here; the 0.43 result is quite abstract.}

\begin{table*}[t]
	\centering
	\caption{$F_1$-measure for tested classifiers on OSC}
	\label{tab:classifier}
	\begin{tabular}{lrr}    \toprule	
		\textbf{Classifier}                   & \textsc{r} & \textsc{qp} \\    \midrule
		\textit{Random Baseline}				 & 0.497869 & 0.486923 \\\midrule
		Decision Trees                & 0.695591 & \underline{0.651462} \\
		Linear SVM                   & \underline{0.705755} & 0.593846 \\
		Naive Bayes                  & 0.661239 & 0.587692 \\ 
		Nearest Neighbors            & 0.701587 & 0.635923 \\   \bottomrule
	\end{tabular}
\end{table*}
\begin{comment}

@@

In this paper we propose methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future. Such methods have applications for synchronization, caching, and more besides. More specifically, we propose a framework based on classifiers and regression that accept features extracted from the queries, from past versions of the data, or from the combination of both. Considering this framework, we propose that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.

Our goal then is to explore this trade-off experimentally. For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc. Settling on Wikidata, we collect 23 versions and curate 221 user-generated queries from which we compute results, statistics and ultimately features that form the gold standard for the tasks. Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task. We make this gold standard available at \ah{...} -- including queries, results and features -- so that it can be reused in further research.

Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for predictions in the OSC and TTL tasks. For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions. Our results show that although features based on historical changes to query results perform best ($F_1 = 0.706$), considering static query features and predicate dynamics alone is competitive ($F_1 =0.651$) when historical results are not available or are prohibitively costly to compute.

In terms of future work, gold standards based on datasets other than Wikidata could be pursued. Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; as previously mentioned, a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features). Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments. Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the proposed predictions.

\ah{Some reference can be shortened to just expand the acronym, no more, for example; \textit{Advances in Information Retrieval -- 34th European Conference on IR Research, ECIR, 2012}; can become \textit{European Conference on IR Research, ECIR, 2012}. Addresses and dates are not necessary (keep the year of course). The workshop name is enough; no need to also give the conference. Keeping the publisher for conference is good, but maybe IEEE Computer Society can become IEEE.}
